Gammapy: A Python package for gamma-ray astronomy

### 2023-05-24 src/ms-review.tex
Add run times of the examples

\subsection{1D Analysis}
\label{ssec:1d-analysis}
One of the most common analysis cases in \gammaray astronomy is measuring the
spectrum of a source in a given region defined on the sky, in conventional
astronomy also called \textit{aperture photometry}. The spectrum is typically measured
in two steps: first a parametric spectral model is fitted to the data and
secondly flux points are computed in a pre-defined set of energy bins. The
result of such an analysis performed on three simulated CTA observations is
shown in Figure~\ref{fig:cta_galactic_center}. In this case the spectrum was
measured in a circular aperture centered on the Galactic Center, in
\gammaray~astronomy often called \enquote{on region}. For such analysis the user first
chooses a region of interest and energy binning, both defined by a
\code{RegionGeom}. In a second step, the events and the IRFs are binned
into maps of this geometry, by the \code{SpectrumDatasetMaker}. All the data and
reduced IRFs are bundled into a \code{SpectrumDataset}. To estimate
the expected background in the \enquote{on region} a \enquote{reflected regions} background
method was used~\citep{Berge07}, represented in \gammapy by the
\code{ReflectedRegionsBackgroundMaker} class. The resulting reflected regions are
illustrated for all three observations overlayed on the counts map in Figure~\ref{fig:cta_galactic_center}.
After reduction, the data were modelled using a forward-folding method and assuming
a point source with a power law spectral shape. The model was defined, using
the \code{SkyModel} class with a \code{PowerLawSpectralModel} spectral component only.
This model was then combined with the \code{SpectrumDataset}, which contains the reduced data 
and fitted using the \code{Fit} class. Based on this best-fit model, the final flux points and corresponding
log-likelihood profiles were computed using the \code{FluxPointsEstimator}.

@@

As the example above shows, the 3D analysis allows to characterize the
morphology of the emission and fit it together with the spectral properties of
the source.  Among the advantages that this provides is the ability to
disentangle the contribution from overlapping sources to the same spatial
region. To highlight this, we define a circular \code{RegionGeom} of radius
0.5\textdegree~ centered around the position of the point source, which is drawn
in the left panel of Figure~\ref{fig:cube_analysis}. We can now compare the
measured excess counts integrated in that region to the expected relative
contribution from each of the three source models. The result can be seen in the right
panel of Figure~\ref{fig:cube_analysis}.

Note that all the models fitted also have a spectral component, from which flux
points can be derived in a similar way as described in Section~\ref{ssec:1d-analysis}.
%\end{figure*}%	\caption{Fermi-LAT TS map in two energy bands} \label{fig:fermi_ts_map}%	\includegraphics[width=1.\textwidth]{figures/fermi_ts_map.pdf}%Ref:~\citep{Stewart2009} \begin{figure*}[t] \centering%\todo{What to do with } Figure~\ref{fig:fermi_ts_map} ?%

\subsection{Temporal Analysis}
\label{ssec:temporal-analysis}
A common use case in many astrophysical scenarios is to study the temporal
variability of a source. The most basic way to do this is to construct a
light curve, i.e., the flux of a source in each given time bin. In \gammapy, this
is done by using the \code{LightCurveEstimator} that fits the normalisation of a
source in each time (and optionally energy) band per observation, keeping constant
other parameters.
For custom time binning, an observation needs to be split into finer time bins using
the \code{Observation.select\_time} method. Figure~\ref{fig:hess_lightcurve_pks}
shows the light curve of the blazar PKS~2155-304 in different energy bands as
observed by the \hess telescope during an exceptional flare on the night of
July 29 - 30, 2006~\cite{2009A&A...502..749A}. The data are publicly available 
as a part of the HESS-DL3-DR1~\cite{HESS_DR1}. Each observation is first split into 10 min smaller
observations, and spectra extracted for each of these within a 0.11\textdegree~radius
around the source. A \code{PowerLawSpectralModel} is fit to all the datasets, leading
to a reconstructed index of $3.54 \pm 0.02$. With this adjusted spectral model
the \code{LightCurveEstimator} runs directly for two energy bands,
\SI{0.5}{TeV}~to~\SI{1.5}{TeV} and \SI{1.5}{TeV}~to~\SI{20}{TeV} respectively.
%
\begin{figure*}[t]
    \sidecaption
	\includegraphics[width=0.6666\textwidth]{figures/hess_lightcurve_pks.pdf}
	\caption{
        Binned PKS~2155-304 light curve in two different energy bands
        as observed by the \hess telescopes in 2006. The coloured markers
        show the flux points in the different energy bands:
		the range from (\SI{0.5}{TeV}~to~\SI{1.5}{TeV} is shown in blue, while
		the range from \SI{1.5}{TeV}~to~\SI{20}{TeV}) is shown in orange.
		The horizontal error illustrates the width of the time bin of 10~min. The vertical
        error bars show the associated asymmetrical flux errors. The marker
        is set to the center of the time bin.
    }
    \label{fig:hess_lightcurve_pks}
\end{figure*}
%
The obtained flux points can be analytically modelled using the available or
user-implemented temporal models. Alternatively, instead of  extracting a
light curve, it is also possible to directly fit temporal models to the reduced
datasets. By associating an appropriate \code{SkyModel}, consisting of both temporal
and spectral components, or using custom temporal models with spectroscopic
variability, to each dataset, a joint fit across the datasets will directly
return the best fit temporal and spectral parameters.

@@

The result of the joint fit is displayed in
Figure~\ref{fig:multi_instrument_analysis}. We remark that the objective of this
exercise is illustrative. We display the flexibility of \gammapy in
simultaneously fitting multi-instrument data even at different levels of
reduction, without aiming to provide a new measurement of the Crab Nebula
spectrum.

@@

As an example application, we use the same multi-instrument
dataset of the Crab Nebula, described in the previous section,
and we apply an inverse
Compton model computed with Naima and wrapped in the \gammapy models
through the \code{NaimaSpectralModel} class. We describe the gamma-ray emission 
with an inverse Compton scenario, considering a log-parabolic
electron distribution that scatters photons from:
%
\begin{itemize}
	\item the synchrotron radiation produced by
	the very same electrons
	\item near and far infrared photon fields
	\item and the cosmic microwave background (CMB)
\end{itemize}
%
We adopt the prescription on the target photon fields provided in the documentation of the \textit{Naima}
package\footnote{\url{https://naima.readthedocs.io/en/stable/examples.html\#crab-nebula-ssc-model}}.
The best-fit inverse Compton spectrum is represented with a red dashed line in
Figure~\ref{fig:multi_instrument_analysis}.

>>>>

\subsection{1D Analysis}
\label{ssec:1d-analysis}
One of the most common analysis cases in \gammaray astronomy is measuring the
spectrum of a source in a given region defined on the sky, in conventional
astronomy also called \textit{aperture photometry}. The spectrum is typically measured
in two steps: first a parametric spectral model is fitted to the data and
secondly flux points are computed in a pre-defined set of energy bins. The
result of such an analysis performed on three simulated CTA observations is
shown in Figure~\ref{fig:cta_galactic_center}. In this case the spectrum was
measured in a circular aperture centered on the Galactic Center, in
\gammaray~astronomy often called \enquote{on region}. For such analysis the user first
chooses a region of interest and energy binning, both defined by a
\code{RegionGeom}. In a second step, the events and the IRFs are binned
into maps of this geometry, by the \code{SpectrumDatasetMaker}. All the data and
reduced IRFs are bundled into a \code{SpectrumDataset}. To estimate
the expected background in the \enquote{on region} a \enquote{reflected regions} background
method was used~\citep{Berge07}, represented in \gammapy by the
\code{ReflectedRegionsBackgroundMaker} class. The resulting reflected regions are
illustrated for all three observations overlayed on the counts map in Figure~\ref{fig:cta_galactic_center}.
After reduction, the data were modelled using a forward-folding method and assuming
a point source with a power law spectral shape. The model was defined, using
the \code{SkyModel} class with a \code{PowerLawSpectralModel} spectral component only.
This model was then combined with the \code{SpectrumDataset}, which contains the reduced data 
and fitted using the \code{Fit} class. Based on this best-fit model, the final flux points and corresponding
log-likelihood profiles were computed using the \code{FluxPointsEstimator}. The
example takes $<10$ seconds to run on a standard laptop with M1 arm64 CPU.

@@

As the example above shows, the 3D analysis allows to characterize the
morphology of the emission and fit it together with the spectral properties of
the source.  Among the advantages that this provides is the ability to
disentangle the contribution from overlapping sources to the same spatial
region. To highlight this, we define a circular \code{RegionGeom} of radius
0.5\textdegree~ centered around the position of the point source, which is drawn
in the left panel of Figure~\ref{fig:cube_analysis}. We can now compare the
measured excess counts integrated in that region to the expected relative
contribution from each of the three source models. The result can be seen in the right
panel of Figure~\ref{fig:cube_analysis}. Note that all the models fitted also have a
spectral component, from which flux points can be derived in a similar way as described
in Section~\ref{ssec:1d-analysis}. The whole example takes $<2$ minutes to run on a standard
laptop with M1 arm64 CPU.
%\end{figure*}%	\caption{Fermi-LAT TS map in two energy bands} \label{fig:fermi_ts_map}%	\includegraphics[width=1.\textwidth]{figures/fermi_ts_map.pdf}%Ref:~\citep{Stewart2009} \begin{figure*}[t] \centering%\todo{What to do with } Figure~\ref{fig:fermi_ts_map} ?%

\subsection{Temporal Analysis}
\label{ssec:temporal-analysis}
A common use case in many astrophysical scenarios is to study the temporal
variability of a source. The most basic way to do this is to construct a
light curve, i.e., the flux of a source in each given time bin. In \gammapy, this
is done by using the \code{LightCurveEstimator} that fits the normalisation of a
source in each time (and optionally energy) band per observation, keeping constant
other parameters.
For custom time binning, an observation needs to be split into finer time bins using
the \code{Observation.select\_time} method. Figure~\ref{fig:hess_lightcurve_pks}
shows the light curve of the blazar PKS~2155-304 in different energy bands as
observed by the \hess telescope during an exceptional flare on the night of
July 29 - 30, 2006~\cite{2009A&A...502..749A}. The data are publicly available 
as a part of the HESS-DL3-DR1~\cite{HESS_DR1}. Each observation is first split into 10 min smaller
observations, and spectra extracted for each of these within a 0.11\textdegree~radius
around the source. A \code{PowerLawSpectralModel} is fit to all the datasets, leading
to a reconstructed index of $3.54 \pm 0.02$. With this adjusted spectral model
the \code{LightCurveEstimator} runs directly for two energy bands,
\SI{0.5}{TeV}~to~\SI{1.5}{TeV} and \SI{1.5}{TeV}~to~\SI{20}{TeV} respectively.
%
\begin{figure*}[t]
    \sidecaption
	\includegraphics[width=0.6666\textwidth]{figures/hess_lightcurve_pks.pdf}
	\caption{
        Binned PKS~2155-304 light curve in two different energy bands
        as observed by the \hess telescopes in 2006. The coloured markers
        show the flux points in the different energy bands:
		the range from (\SI{0.5}{TeV}~to~\SI{1.5}{TeV} is shown in blue, while
		the range from \SI{1.5}{TeV}~to~\SI{20}{TeV}) is shown in orange.
		The horizontal error illustrates the width of the time bin of 10~min. The vertical
        error bars show the associated asymmetrical flux errors. The marker
        is set to the center of the time bin.
    }
    \label{fig:hess_lightcurve_pks}
\end{figure*}
%
The obtained flux points can be analytically modelled using the available or
user-implemented temporal models. Alternatively, instead of  extracting a
light curve, it is also possible to directly fit temporal models to the reduced
datasets. By associating an appropriate \code{SkyModel}, consisting of both temporal
and spectral components, or using custom temporal models with spectroscopic
variability, to each dataset, a joint fit across the datasets will directly
return the best fit temporal and spectral parameters. The light curve data reduction
and computation of flux points takes about $0.5$ minutes to run on a standard laptop with M1 arm64 CPU.

@@

The result of the joint fit is displayed in
Figure~\ref{fig:multi_instrument_analysis}. We remark that the objective of this
exercise is illustrative. We display the flexibility of \gammapy in
simultaneously fitting multi-instrument data even at different levels of
reduction, without aiming to provide a new measurement of the Crab Nebula
spectrum. The spectral fit takes $<10$ seconds to run on a standard laptop with M1 arm64 CPU.

@@

As an example application, we use the same multi-instrument
dataset of the Crab Nebula, described in the previous section,
and we apply an inverse
Compton model computed with Naima and wrapped in the \gammapy models
through the \code{NaimaSpectralModel} class. We describe the gamma-ray emission 
with an inverse Compton scenario, considering a log-parabolic
electron distribution that scatters photons from:
%
\begin{itemize}
	\item the synchrotron radiation produced by
	the very same electrons
	\item near and far infrared photon fields
	\item and the cosmic microwave background (CMB)
\end{itemize}
%
We adopt the prescription on the target photon fields provided in the documentation of the \textit{Naima}
package\footnote{\url{https://naima.readthedocs.io/en/stable/examples.html\#crab-nebula-ssc-model}}.
The best-fit inverse Compton spectrum is represented with a red dashed line in
Figure~\ref{fig:multi_instrument_analysis}. The fit of the astrophysical model
takes $<5$ minutes to run on a standard laptop with M1 arm64 CPU.
